Bird beak's elaboration and innovation

### 2022-07-19 Analysis/03.1-Elaboration_innovation_analysis[statistics].Rmd
calcualting the stats for all ellipses

```{r, eval = TRUE, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, print = FALSE}
library(dispRity)
source("../Functions/ploting.utilities.R")
```

@@

```{r, eval = TRUE, echo = FALSE}
## Load the data for all birds
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")
## Select the data
shapespace <- shapespace_allbirds_lvl_superorder_order[[1]]$space
tree <- shapespace_allbirds_lvl_superorder_order[[1]]$consensus_tree
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Load the covar
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
shapespace_allbirds_lvl_superorder_order <- shapespace_allbirds_lvl_superorder_order[[1]]

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order) == "animal")
super_lvl <- grep("level1_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))
order_lvl <- grep("level2_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))

## Create the subset for all birds
all_birds_shapespace <- MCMCglmm.subsets(n = 500, # Just run it on 500 matrices for now
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(super_lvl, order_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level1)[-1], levels(shapespace$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

# Load the data for the passeriformes
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")
## Select the data
shapespace <- shapespace_allbirds_lvl_superorder_order[[1]]$space
tree <- shapespace_allbirds_lvl_superorder_order[[1]]$consensus_tree
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Load the covar
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
shapespace_allbirds_lvl_superorder_order <- shapespace_allbirds_lvl_superorder_order[[1]]

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order) == "animal")
super_lvl <- grep("level1_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))
order_lvl <- grep("level2_", MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order))

## Create the subset for all birds
all_birds_shapespace <- MCMCglmm.subsets(n = 500, # Just run it on 500 matrices for now
      data          = shapespace,
      posteriors    = shapespace_allbirds_lvl_superorder_order,
      group         = MCMCglmm.levels(shapespace_allbirds_lvl_superorder_order)[c(super_lvl, order_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace$level1)[-1], levels(shapespace$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

```

@@

```{r, eval = FALSE, echo = FALSE}
source("../Functions/stats.utilities.R")
## General measurement
super_orders_stats <- ellipse.stats(superorder_shapespace)
orders_stats <- ellipse.stats(order_shapespace)
suborder_stats <- ellipse.stats(suborder_shapespace)
family_stats <- ellipse.stats(family_shapespace)

ellipse_stats <- list(super_orders_stats = super_orders_stats, 
                      orders_stats = orders_stats,
                      suborder_stats = suborder_stats,
                      family_stats = family_stats)
save(ellipse_stats, file = "../Data/Processed/ellipse_stats.rda")
```

>>>>

```{r, eval = TRUE, echo = FALSE, results = 'hide', message = FALSE, warning = FALSE, print = FALSE}
library(dispRity)
```

@@

```{r, eval = TRUE, echo = FALSE}
## Load the data for all birds
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")
## Select the data
shapespace_allbirds <- shapespace_allbirds_lvl_superorder_order[[1]]$space
## Clean garbage
rm(shapespace_allbirds_lvl_superorder_order)

## Load the covar
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.MCMCglmm")
shapespace_allbirds_posteriors <- shapespace_allbirds_lvl_superorder_order[[1]]
rm(shapespace_allbirds_lvl_superorder_order)

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_allbirds_posteriors) == "animal")
super_lvl <- grep("level1_", MCMCglmm.levels(shapespace_allbirds_posteriors))
order_lvl <- grep("level2_", MCMCglmm.levels(shapespace_allbirds_posteriors))

## Create the subset for all birds
all_birds_shapespace <- MCMCglmm.subsets(n = 500, # Just run it on 500 matrices for now
      data          = shapespace_allbirds,
      posteriors    = shapespace_allbirds_posteriors,
      group         = MCMCglmm.levels(shapespace_allbirds_posteriors)[c(super_lvl, order_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace_allbirds$level1)[-1], levels(shapespace_allbirds$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_allbirds_posteriors)

## Load the covar
load(file = "../Data/Processed/shapespace_passeriformes.MCMCglmm")
shapespace_passeriformes_posteriors <- shapespace_passeriformes[[1]]
rm(shapespace_passeriformes)
## Load the data for the passeriformes
load(file = "../Data/Processed/shapespace_passeriformes_two_levels.rda")
## Select the data
shapespace_passeriformes <- shapespace_passeriformes_two_levels[[1]]$space
## Clean garbage
rm(shapespace_passeriformes_two_levels)

## Get the levels
phylo_lvl <- which(MCMCglmm.levels(shapespace_passeriformes_posteriors) == "animal")
suborder_lvl <- grep("level1_", MCMCglmm.levels(shapespace_passeriformes_posteriors))
family_lvl <- grep("level2_", MCMCglmm.levels(shapespace_passeriformes_posteriors))

## Create the subset for all birds
passeriformes_shapespace <- MCMCglmm.subsets(n = 500, # Just run it on 500 matrices for now
      data          = shapespace_passeriformes,
      posteriors    = shapespace_passeriformes_posteriors,
      group         = MCMCglmm.levels(shapespace_passeriformes_posteriors)[c(suborder_lvl, family_lvl, phylo_lvl)],
      rename.groups = c(levels(shapespace_passeriformes$level1)[-1], levels(shapespace_passeriformes$level2)[-1], "phylogeny")) #[-1] is because the first level is ""
## Clean garbage
rm(shapespace_passeriformes_posteriors)
```

```{r}
## Creating the nested levels subgroups
make.nested.subset <- function(group, data, group.data) {
    ## Get the subsets within the group
    level2_groups <- as.character(unique(group.data$level2[group.data$level1 == group]))
    if(any(empty <- level2_groups == "")) {
        level2_groups <- level2_groups[-which(empty)]
    }
    if(length(level2_groups) < 1) {
        return(NULL)
    }
    ## Create a subset
    return(get.subsets(data, subsets = c(level2_groups, group)))
}

## Get the subgroups for all_birds
all_birds_shapespace_subgroups <- lapply(as.list(levels(shapespace_allbirds$level1)[-1]), make.nested.subset, data = all_birds_shapespace, group.data = shapespace_allbirds)
names(all_birds_shapespace_subgroups) <- levels(shapespace_allbirds$level1)[-1]
## Get the subgroups for the passeriformes
passeriformes_shapespace_subgroups <- lapply(as.list(levels(shapespace_passeriformes$level1)[-1]), make.nested.subset, data = passeriformes_shapespace, group.data = shapespace_passeriformes)
names(passeriformes_shapespace_subgroups) <- levels(shapespace_passeriformes$level1)[-1]
```

@@

```{r, eval = FALSE, echo = FALSE}
source("../Functions/stats.utilities.R")

## Fucked up thing happening here:
cd("~/Packaging/dispRity/")
load_all()

## General measurement
all_birds_all_levels <- ellipse.stats(all_birds_shapespace)
passeriformes_all_levels <- ellipse.stats(passeriformes_shapespace)

## Nested measurements
all_birds_nested_levels <- lapply(all_birds_shapespace_subgroups, ellipse.stats)
passeriformes_nested_levels <- lapply(passeriformes_shapespace_subgroups, ellipse.stats)

## Big measurements list
ellipses_stats <- list("all_birds" = all_birds_all_levels)
ellipses_stats <- c(ellipses_stats, all_birds_nested_levels)
ellipses_stats$Passeriformes <- passeriformes_all_levels
ellipses_stats <- c(ellipses_stats, passeriformes_nested_levels)

save(ellipses_stats, file = "../Data/Processed/ellipses_stats.rda")
```
